---
title: "P8105 HOMEWORK1"
author: "Junhui Mi"
date: 2019-09-14
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

The purpose of this homework is to work with data and present plots using _ggplot_.

## Problem 1

First I create a dataframe _df_1_ containing variables for problem 1.

```{r df_1_create}
df_1 = tibble(
  norm_samp = rnorm(8),
  norm_samp_pos = norm_samp > 0,
  chara = c("A", "B", "C", "D", "E", "F", "G", "H"),
  fact = factor(c("data", "biostat", "epi", "biostat", "data",
                  "biostat","data","data"), 
                levels = c("data", "epi", "biostat"))
)
```

Then I take the mean of each variable in my dataframe to check if it works.

```{r df_1_mean}
mean(pull(df_1, norm_samp))
mean(pull(df_1, norm_samp_pos))
mean(pull(df_1, chara))
mean(pull(df_1, fact))
```

I find that mean function works for numeric and logical variables but not for character and factor variables.

I try to apply "as.numeric" function to change logical, character, and factor variables into numeric variables.

```{r df_1_numeric, results = 'hide', warning = FALSE}
as.numeric(pull(df_1, norm_samp_pos))
as.numeric(pull(df_1, chara))
as.numeric(pull(df_1, fact))
```

The results show that logic variables can be changed to numeric variables 0 and 1. However, character and factor variables can not be changed directly because their values are unknown. It counts for the mean function not working on character and factor variables.

Finally, I do calculation using those variables.

```{r df_1_computation}
logic_num <- as.numeric(pull(df_1, norm_samp_pos)) * pull(df_1, norm_samp)
logic_num
logic_fact <- as.factor(pull(df_1, norm_samp_pos)) * pull(df_1, norm_samp)
logic_fact
logic_fact_num <- as.numeric(as.factor(pull(df_1, norm_samp_pos))) * 
  pull(df_1, norm_samp)
logic_fact_num
```

It shows that numeric variables can do calculation while factor variables can not.

## Problem 2

First I create a dataframe _df_2_ containing variables for my plots.

```{r df_2_create, eval = TRUE}
df_2 = tibble(
  x = rnorm(500),
  y = rnorm(500),
  logi = x + y > 1,
  num = as.numeric(logi),
  fac = as.factor(logi)
)
```

The size of the dataset is `r nrow(df_2)` rows and `r ncol(df_2)` columns.

The mean of x is `r mean(pull(df_2, x))`, the median of x is `r median(pull(df_2, x))`, and the  standard deviation of x is `r sd(pull(df_2, x))`.

The proportion of cases for which x + y > 1 is `r sum(pull(df_2, num))/500`

Then I show three scatterplots of _y_ vs _x_. 

_Plot 1_

```{r yx_scatter_1}
ggplot(df_2, aes(x = x, y = y, color = pull(df_2, logi))) + geom_point()
ggsave("my first scatterplot for HW1.png")
```

_Plot 2_
```{r yx_scatter_2}
ggplot(df_2, aes(x = x, y = y, color = pull(df_2, num))) + geom_point()
```

_Plot 3_
```{r yx_scatter_3}
ggplot(df_2, aes(x = x, y = y, color = pull(df_2, fac))) + geom_point()
```

The results above shows that:

**(1)** For all three plots, arrays are divided into two groups: 
one group is "x + y > 1" and the other is "x + y <= 1"

**(2)** ggplot has its color scale for logical variables "TRUE" and "FALSE", which is shown as pink dots and green dots in Plot 1.

**(3)** ggplot has a gradient color between 0 and 1, from dark blue to light blue, as is shown in Plot 2.

**(4)** If factor variables only contain levels of "TRUE" and "FALSE", then ggplot color scale is same as logical variables, as is shown in Plot 3.
